---
title: "p8105_hw2_yw3095"
author: "Yixuan Wang"
date: "September 30, 2018"
output: github_document
---

```{r setup, include=FALSE}
getwd()
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
```
##Problem 1
This problem focuses on NYC Transit data

*   Read and clean the data
```{r}
hw2_transit = 
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv", 
           col_names = TRUE) %>% 
  janitor::clean_names() %>% 
  gather(key = route_number, 
         value = route_name, 
         route1:route11) %>% 
  select(line, 
         station_name,
         station_latitude,
         station_longitude,
         route_number, 
         route_name, 
         entry, 
         vending, 
         entrance_type, 
         ada) %>% 
  mutate(entry = ifelse(entry == "YES", yes = TRUE, no = FALSE))
head(hw2_transit)
```
**steps to clean the data**

describe what happened

**questions**

* How many distinct stations are there?

    * There are `r count(distinct(hw2_transit, line, station_name))` distinct stations.

* How many distinct stations are ADA compliant?

    * There are `r count(filter(hw2_transit, ada == TRUE) %>% distinct(.data., line, station_name))` distinct stations are ADA compliant.

* What proportion of station entrances / exits without vending allow entrance?

    *   `r round (count(filter(hw2_transit, ada == TRUE) %>% distinct(.data., line, station_name))/count(distinct(hw2_transit, line, station_name))*100, 2)`% of station entrances / exits without vending allow entrance.

##Problem 2
This problem focuses on the Mr. Trash Wheel dataset.

*   Read and clean the data  total!!!
```{r}
hw2_wheel = 
  read_excel("./data/HealthyHarborWaterWheelTotals2017-9-26.xlsx", 
             range = cell_cols(1:14),
             col_names = TRUE) %>% 
  janitor::clean_names() %>% 
  filter(date != 0) %>% 
  mutate(sports_balls = as.integer(round(sports_balls)))
hw2_wheel
```
*   Read and clean precipitation data for 2016
```{r}
pre_2016 = 
  read_excel("./data/HealthyHarborWaterWheelTotals2017-9-26.xlsx", 
                      sheet = 4,
                      range = cell_rows(2:14),
                      col_names = TRUE) %>%
  janitor::clean_names() %>% 
  mutate(year = 2016)
pre_2016
```
*   Read and clean precipitation data for 2017
```{r}
pre_2017 = 
  read_excel("./data/HealthyHarborWaterWheelTotals2017-9-26.xlsx", 
                      sheet = 3,
                      range = cell_rows(2:10),
                      col_names = TRUE) %>%
  janitor::clean_names() %>% 
  mutate(year = 2017)
pre_2017
```
